---
title: "Statistical report: SGLT2i across the EF spectrum"
subtitle: Draft
date: "`r Sys.Date()`"
cover: static/ki_logo_vert_rgb.png
editor: source
format:
  pdf:
    documentclass: scrartcl
    template-partials:
      - "static/before-body.tex"
      - "static/_titlepage.tex"
    include-in-header: 
      - "static/in-header.tex"
    toc: true
    toc-depth: 3
    toc-title: Contents
    lof: true
    lot: true
    cap-location: top
    number-sections: true
    colorlinks: false
    keep-tex: false
bibliography: static/references.bib
csl: static/bmj.csl
link-citations: true
link-bibliography: true
nocite: '@*'
knitr:
  opts_chunk: 
    dev: cairo_pdf
    collapse: true
    comment: "" 
    echo: FALSE
    include: TRUE
    warning: FALSE
    message: FALSE
    fig.pos: "H"
    fig.path: "../output/figs/"
    fig.height: 6
    fig.width: 10
    R.options:
      knitr.graphics.auto_pdf: true
---

```{r}
#| label: set-up-load-data
options(knitr.kable.NA = "")

# load packages, globals and project specific functions
source(here::here("setup/setup.R"))

# load data
load(here("data/clean-data/rsdata.RData"))
load(here("data/clean-data/nprdata.RData"))

# load workbook to write tables to Excel
wb <- loadWorkbook(here("output/tabs/tables.xlsx"))
sheets <- names(wb)

# load pptx file with figs
figs <- officer::read_pptx(path = here::here("output/figs/figs.pptx"))
pptx <- FALSE
```           

\newpage

# Data

## Data sources

The Swedish Heart Failure Registry (SwedeHF) is an ongoing nationwide quality registry started in 2000, 
that includes in- and out-hospital patients with HF, regardless of EF and previous history of HF[@swedehf]. 
Around 80 variables are recorded at 
discharge from hospital or at the outpatient visit, i.e. index date, and 
entered into an electronic database managed by the Uppsala Clinical Research 
Center (Uppsala, Sweden). Up to April 2017 the only inclusion criterion was a 
clinical diagnosis of HF, which was thereafter defined according to the ICD-10 
codes I50.0, I50.1, I50.9, I42.0, I42.6, I42.7, I25.5, I11.0, I13.0, I13.2. 
In 2021, 69 of 76 Swedish hospitals enrolled patients in SwedeHF, which had a 32% coverage of the prevalent HF population in Sweden[@annualreport]. 

SwedeHF was linked with national Swedish registries (SHFDB version 4.2.0) through the 
Swedish personal identity number (PIN)[@pin]: 

- The National Patient Register (The National Board of Health and Welfare) for definition of the NPR population and
additional comorbidities
- The Cause of Death Register[@dors] (The National Board of Health and Welfare) for outcomes 
death (used for censoring)
- The National Prescribed Drug Register (The National Board of Health and Welfare) for SGLT2i dispensations
- Longitudinal integrated database for health insurance and labour market studies (LISA)[@lisa] 
and the Register of the Total Population (Statistics Sweden) for socio-economic factors

## Acknowledgments

Please include "We thank all staff members at care units reporting into the SwedeHF registry for their contribution." 
in the Acknowledgements of any resulting articles. 

## Ethics and Informed consent

The study was approved by the Swedish Ethical Review Authority, dnr 2021-04326. 
Individual consent was not required, but patients were informed of entry into SwedeHF and able to opt‐out.

## Inclusion/exclusion criteria

The analyses were performed for 2 populations: 

### SwedeHF population

```{r}
#| label: tbl-flow-rs
#| tbl-cap: Information for flowchart - SwedeHF
#| tbl-pos: "H"

default_kable(flow_rs) %>%
  row_spec(c(1, 11), bold = T)
```

First patient in: `r min(rsdata$shf_indexdtm)` and last patient in: `r max(rsdata$shf_indexdtm)`. 

Baseline for this population is date of visits or date of discharge for the visit/hospitalization recorded in SwedeHF. 

### NPR population

Patients with $\geq2$ ICD-10 code for HF in main position in the National Patient Register 
(ICD-10: I110, I130, I132, I255, I420, I423, 1425-9, I43, I50, J81, K761) 
starting from 2000-01-01 with no previous ICD-9/10 code for HF 
(I110 I130 I132 I255 I420 I423 1425 I426 I427 I428 I429 I43 I50 J81 K761 ICD-9: 414W 425E 425F 425G 425H 425W 425X 428) 
since 1987 were defined as the NPR population. 

```{r}
#| label: tbl-flow-npr
#| tbl-cap: Information for flowchart - NPR
#| tbl-pos: "H"

default_kable(flow_npr)
```

First patient in: `r min(nprdata$indexdtm)` and last patient in: `r max(nprdata$indexdtm)`.  

Baseline for this population is date of visits or date of discharge for the second visit/hospitalization including a hospitalization in the NPR. 

The NPR population was only used to present trends over time. 

#### Limitation 

Note that the NPR population is included up until 2021-12-31 so newly diagnosed patients are not included in the population. 

## Definitions

Information on data sources, definitions etc. are found https://kiheartfailure.github.io/shfdb4/.

{{< include src/vars.qmd >}}

Please CHECK ATC codes for sglt2!

# Statistical analysis 

## General

All analyses were performed using `r sessionInfo()$R.version$version.string` [@r]. 
The level of significance is set to 5%, two-sided. No adjustment for multiple 
comparisons were made and therefore the results should be viewed with care.

## Missing data

Missing data was imputed with multiple imputation (n = 10) using mice [@mice] 
and Rubin’s rules was used for combining estimates and standard errors across imputed data sets. 
Variables included in the model are indicated in @tbl-base-ref - @tbl-base-pef. 

## Subgroups

All analyses from SwedeHF are presented by EF defined as HFrEF (LVEF <40%), HFrEF (LVEF 40-49%) and HFpEF (LVEF $\geq$ 50%). 

Further, on the SwedeHF population, selected analyses are performed by `r paste0(subgroupvars_rs$label, collapse = ", ")`. 

## Baseline characteristics

Baseline characteristics are presented by SGLT2i. 
Categorical variables are presented with n and percentage and tested for 
differences between groups with the chi-square test. 
Continuous variables are presented with median [first quartile-third quartile] and tested for 
differences between groups with the Kruskal-Wallis test. 

## Trends in SGLT2i over time

Trends in SGLT2i prevalent use, incident use and discontinuation was presented over calender time. 

### Prevalent use

The denominator was defined as patients not censored (alive and not emigrated from Sweden) and available in the population, 
for the NPR population all patients since inclusion stopped 2021-12-31 and for the SwedeHF population patients recorded in SwedeHF at the time. 

The numerator was defined as patients in the denominator with an SGLT2i dispensation within 4 months prior. 

### Incident use

Incident use was defined as prevalent use with the exception that only patients without a SGLT2i dispensation 16-4 months prior to baseline were included (thereby considered as SGLT2i naive).

### Discontinuation

Discontinuation was defined similar to the above but only patients with a SGLT2i dispensation 8-4 months prior to baseline were included (thereby considered to be SGLT2i users).

## Associations between baseline characteristics and SGLT2i

The association between baseline characteristics and SGLT2i was evaluated using a multivariable logistic regression 
including the variables shown in @tbl-predictors and @fig-predictors. The variables were selected based on clinical relevance. 

### Assumptions

TO CHECK: Possible outliers were evaluated using Cook's distance and multicollinearity was 
evaluated using the Variance Inflation Factor (VIF) without any cause for action. 

# Results

```{r}
med <- rsdata %>%
  summarise(
    med = fn(median(shf_age), dig = 0),
    q1 = fn(quantile(shf_age, probs = 0.25), dig = 0),
    q3 = fn(quantile(shf_age, probs = 0.75), dig = 0)
  ) %>%
  mutate(out = paste0(med, " (", q1, "-", q3, ")")) %>%
  pull(out)

percfem <- rsdata %>%
  count(shf_sex) %>%
  mutate(perc = fn(n / sum(n) * 100, 0)) %>%
  filter(shf_sex == "Female") %>%
  pull(perc)

med_npr <- nprdata %>%
  summarise(
    med = fn(median(shf_age), dig = 0),
    q1 = fn(quantile(shf_age, probs = 0.25), dig = 0),
    q3 = fn(quantile(shf_age, probs = 0.75), dig = 0)
  ) %>%
  mutate(out = paste0(med, " (", q1, "-", q3, ")")) %>%
  pull(out)

percfem_npr <- nprdata %>%
  count(shf_sex) %>%
  mutate(perc = fn(n / sum(n) * 100, 0)) %>%
  filter(shf_sex == "Female") %>%
  pull(perc)
```

The median (q1-q3) age is `r med` and `r percfem`% females in the SwedeHF population and `r med_npr` and `r percfem_npr`% in the NPR population.    

## Baseline characteristics

{{< include src/base_tab.qmd >}}

\clearpage

{{< include src/use_geo.qmd >}}

## Trends in SGLT2i over time

### SwedeHF population

{{< include src/overtime_figs_rs.qmd >}}

\clearpage 

### NPR population

{{< include src/overtime_figs_npr.qmd >}}

\clearpage

## Associations between baseline characteristics and SGLT2i

{{< include src/predictors.qmd >}}

\clearpage

# Reproducibility

## R code

The R code for all data handling and statistical analyses are found: 

https://github.com/KIHeartFailure/swedehf-sglt2ef. On publication
the repository will be made public so as to 
link to it from the resulting article for increased transparency and code sharing.
No data or output is stored in the repository. 

## R session information {#sessioninfo}

```{r sessinfo}
sessionInfo()
```

# References

::: {#refs}
:::
