---
title: "Statistical report: SGLT2i across the EF spectrum"
date: "`r Sys.Date()`"
cover: static/ki_logo_vert_rgb.png
editor: source
format:
  pdf:
    documentclass: scrartcl
    template-partials:
      - "static/before-body.tex"
      - "static/_titlepage.tex"
    include-in-header: 
      - "static/in-header.tex"
    toc: true
    toc-depth: 3
    toc-title: Contents
    lof: true
    lot: true
    cap-location: top
    number-sections: true
    colorlinks: false
    keep-tex: false
bibliography: static/references.bib
csl: static/bmj.csl
link-citations: true
link-bibliography: true
nocite: '@*'
knitr:
  opts_chunk: 
    dev: cairo_pdf
    collapse: true
    comment: "" 
    echo: FALSE
    include: TRUE
    warning: FALSE
    message: FALSE
    fig.pos: "H"
    fig.path: "../output/figs/"
    fig.height: 6
    fig.width: 10
    R.options:
      knitr.graphics.auto_pdf: true
---

```{r}
#| label: set-up-load-data
options(knitr.kable.NA = "")

# load packages, globals and project specific functions
source(here::here("setup/setup.R"))

# load data
load(here("data/clean-data/rsdata.RData"))

# load workbook to write tables to Excel
wb <- loadWorkbook(here("output/tabs/tables.xlsx"))
sheets <- names(wb)

# load pptx file with figs
figs <- officer::read_pptx(path = here::here("output/figs/figs.pptx"))
pptx <- T
```           

\newpage

# Data

## Data sources

The Swedish Heart Failure Registry (SwedeHF) is an ongoing nationwide quality registry started in 2000, 
that includes in- and out-hospital patients with HF, regardless of EF and previous history of HF[@swedehf]. 
Around 80 variables are recorded at 
discharge from hospital or at the outpatient visit, i.e. index date, and 
entered into an electronic database managed by the Uppsala Clinical Research 
Center (Uppsala, Sweden). Up to April 2017 the only inclusion criterion was a 
clinical diagnosis of HF, which was thereafter defined according to the ICD-10 
codes I50.0, I50.1, I50.9, I42.0, I42.6, I42.7, I25.5, I11.0, I13.0, I13.2. 
In 2023, 69 of 76 Swedish hospitals enrolled patients in SwedeHF, which had a 59% coverage of the prevalent HF population in Sweden[@annualreport]. 

SwedeHF was linked with national Swedish registries (SHFDB version 4.2.1) through the 
Swedish personal identity number (PIN)[@pin]: 

- The National Patient Register (The National Board of Health and Welfare) for additional comorbidities
- The Cause of Death Register[@dors] (The National Board of Health and Welfare) for outcomes 
death (used for censoring)
- The National Prescribed Drug Register (The National Board of Health and Welfare) for SGLT2i dispensations
- Longitudinal integrated database for health insurance and labour market studies (LISA)[@lisa] 
and the Register of the Total Population (Statistics Sweden) for socio-economic factors

## Acknowledgments

Please include "We thank all staff members at care units reporting into the SwedeHF registry for their contribution." 
in the Acknowledgements of any resulting articles. 

## Ethics and Informed consent

The study was approved by the Swedish Ethical Review Authority, dnr 2021-04326. 
Individual consent was not required, but patients were informed of entry into SwedeHF and able to opt‐out.

## Inclusion/exclusion criteria

```{r}
#| label: tbl-flow-rs
#| tbl-cap: Information for flowchart - SwedeHF
#| tbl-pos: "H"

default_kable(flow) %>%
  row_spec(c(1, 11), bold = T)
```

First patient in: `r min(rsdata$shf_indexdtm)` and last patient in: `r max(rsdata$shf_indexdtm)` 
from approximately `r rsdata %>% distinct(shf_centre) %>% count() %>% pull(n)` enrolling centers. 

Baseline is date of visits or date of discharge for the visit/hospitalization recorded in SwedeHF. 

## Definitions

Information on data sources, definitions etc. are found https://kiheartfailure.github.io/shfdb4/.

{{< include src/vars.qmd >}}

# Statistical analysis 

## General

All analyses were performed using `r sessionInfo()$R.version$version.string` [@r]. 
The level of significance is set to 5%, two-sided. No adjustment for multiple 
comparisons were made and therefore the results should be viewed with care.

## Missing data

Missing data was imputed with multiple imputation (n = 10) using mice [@mice] 
and Rubin’s rules was used for combining estimates and standard errors across imputed data sets. 
Variables included in the model are indicated in @tbl-base-ref - @tbl-base-pef. SGLT2i is not included in the model. 

## Subgroups

All analyses are presented by EF defined as HFrEF (LVEF <40%), HFrEF (LVEF 40-49%) and HFpEF (LVEF $\geq$ 50%). 

## Baseline characteristics

Baseline characteristics are presented by SGLT2i. 
Categorical variables are presented with n and percentage and tested for 
differences between groups with the chi-square test. 
Continuous variables are presented with median [first quartile-third quartile] and tested for 
differences between groups with the Kruskal-Wallis test. 

## Temporal trends of SGLT2i use

Trends in SGLT2i use was presented by inclusion Month/Year in SwedeHF by EF, `r paste0(subgroupvars$label, collapse = ", ")`. 

## Associations between baseline characteristics and SGLT2i

The association between baseline characteristics and SGLT2i was evaluated using univariable and multivariable logistic regression models
including the variables shown in @tbl-predictors and @fig-predictors. The variables were selected based on clinical relevance. 

This was done partly "Overall", and partly by EF, 
where the interaction between EF and the respective variable was included in order to evaluate possible differences across the EF spectrum. 
This was done partly with only EF, the respective variable and the interaction between the two and presented as "Crude" estimates and 
partly including all the other variables (but with only the interaction between EF and the respective variable) and presented as "Adjusted". 

### Assumptions

Possible outliers were evaluated using Cook's distance and multicollinearity was 
evaluated using the Variance Inflation Factor (VIF) without any cause for action. 

## Discontinuation

A patient was assumed to have discontinued if there was no dispensed prescription within a time frame of 5 months from the previous prescription. 
The time of discontinuation was set to the last prescription + 4 months 
and was presented using a cumulative incidence curve where patients were censored 2023-12-31, emigration from Sweden or death.  

\clearpage

# Results

```{r}
med <- rsdata %>%
  summarise(
    med = fn(median(shf_age), dig = 0),
    q1 = fn(quantile(shf_age, probs = 0.25), dig = 0),
    q3 = fn(quantile(shf_age, probs = 0.75), dig = 0)
  ) %>%
  mutate(out = paste0(med, " (", q1, "-", q3, ")")) %>%
  pull(out)

percfem <- rsdata %>%
  count(shf_sex) %>%
  mutate(perc = fn(n / sum(n) * 100, 0)) %>%
  filter(shf_sex == "Female") %>%
  pull(perc)
```

The median (q1-q3) age is `r med` and `r percfem`% females.    

## Baseline characteristics

{{< include src/base_tab.qmd >}}

\clearpage

{{< include src/use_geo.qmd >}}

## Temporal trends of SGLT2i use

{{< include src/overtime_figs.qmd >}}

\clearpage

## Associations between baseline characteristics and SGLT2i

{{< include src/predictors.qmd >}}

\clearpage

## Discontinuation

{{< include src/disc.qmd >}}

\clearpage

# Reproducibility

## R code

The R code for all data handling and statistical analyses are found: 

https://github.com/KIHeartFailure/swedehf-sglt2ef. On publication
the repository will be made public so as to 
link to it from the resulting article for increased transparency and code sharing.
No data or output is stored in the repository. 

## R session information {#sessioninfo}

```{r sessinfo}
sessionInfo()
```

# References

::: {#refs}
:::
